---
title: "tweets_SNA"
author: "Shreyans Kothari"
date: "4/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(igraph)
library(ggnetwork)
library(tidygraph)
library(visNetwork)
```


```{r}
setwd("/Users/shreyanskothari/Desktop/Data Visualization/ETL")
hashtags_2011 = read.csv("hashtags_2011.csv")
hashtags_2021 = read.csv("hashtags_2021.csv")

```

```{r}
hashtags_2011 <- hashtags_2011 %>% select(-"X",-"hashtags", -"split_hash")
hashtags_2021 <- hashtags_2021 %>% select(-"X",-"X.1")
```

```{r}
#2011
hash_2011 = t(as.matrix(hashtags_2011)) %*% as.matrix(hashtags_2011)
x_2011 <- as.matrix(hash_2011)
degree_x_2011 <- graph.adjacency(x_2011, mode ="undirected",diag=F)

#2021
hash_2021 = t(as.matrix(hashtags_2021)) %*% as.matrix(hashtags_2021)
x_2021 <- as.matrix(hash_2021)
degree_x_2021 <- graph.adjacency(x_2021, mode ="undirected", diag = F)

```


```{r}
p1 = plot(network, layout=layout.sphere, main="sphere")
p2 = plot(delete.vertices(simplify(degree_x_2011), degree(simplify(degree_x_2011))==0))
p3 = plot(delete.vertices(simplify(degree_x_2021), degree(simplify(degree_x_2021))==0), layout= layout.auto)
p4 = plot(delete.vertices(simplify(degree_x_2021), degree(simplify(degree_x_2021))==0), layout= layout.auto, vertex.size=10)
p5 = plot(delete.vertices(simplify(degree_x_2011), degree(simplify(degree_x_2011))==0), layout= layout.auto, vertex.size=10)
p6 = plot(delete.vertices(simplify(degree_x_2021), degree(simplify(degree_x_2021))==0), layout= layout_nicely, vertex.size=10)

p7 = plot(delete.vertices(simplify(degree_x_2011), degree(simplify(degree_x_2021))==0), layout= layout_nicely, vertex.size=5, label.size = 2)

p_2011 = plot(delete.vertices(simplify(degree_x_2011), degree(simplify(degree_x_2011))==0), layout= layout.davidson.harel, vertex.size=3, curved = F, asp = 0, main = "2011 Hashtags",vertex.label.cex = 0.8, vertex.label.dist = 0.4, vertex.label.color = "black")

p_2021 = plot(delete.vertices(simplify(degree_x_2021), degree(simplify(degree_x_2021))==0), layout= layout.davidson.harel, vertex.size=3, curved = F, asp = 0, main = "2021 Hashtags",vertex.label.cex = 0.55, vertex.label.dist = 0.6, vertex.label.color = "black")


# 2021 With ALL nodes
plot(simplify(graph_2021), layout= layout.davidson.harel, vertex.size=3, curved = F, asp = 0, main = "2021 Hashtags",vertex.label.cex = 0.55, vertex.label.dist = 0.6, vertex.label.color = "black")


```

# Network without unconnected nodes 
```{r}
graph_2011 <- delete.vertices((degree_x_2011), degree((degree_x_2011))==0)
graph_2021 <- delete.vertices((degree_x_2021), degree((degree_x_2021))==0)

```




# Girvan-Neuman
```{r}


set.seed(12)
# Girvan-Newman community detection algorithm
gn_2011 = edge.betweenness.community(graph_2011, directed = FALSE)

```

```{r}
plot(gn, (graph_2011), layout= layout.davidson.harel, vertex.size=3, curved = F, asp = 0, vertex.label.cex = 0.55, edge.color = rgb(0,0,0,.25), main = "Girvan-Neuman")
```


# Random Walk
```{r}
set.seed(12)
walk_2011 = walktrap.community(graph_2011)
```


```{r}
vis_graph_2011 <- toVisNetworkData(graph_2011)
nodes_2011 <- vis_graph_2011$nodes
edges_2011 <- vis_graph_2011$edges


vis_graph_2021 <- toVisNetworkData(graph_2021)
nodes_2021 <- vis_graph_2021$nodes
edges_2021 <- vis_graph_2021$edges

# OR directly input igraph object to visIgraph()

```

```{r}
visNetwork(vis_graph_2021$nodes, vis_graph_2021$edges) %>% 
  visIgraphLayout(layout = "layout_with_fr") %>% 
  visEdges(arrows = "middle")

visNetwork(nodes_2011, edges_2011)%>% visOptions(selectedBy = "label", 
                                                                                                                  highlightNearest = T, 
                                                                                                                  nodesIdSelection = T)%>%
    visEdges(physics = F, smooth = F) %>% visIgraphLayout(layout = "layout.davidson.harel") %>% visNodes(labelHighlightBold = T, label = nodes_2011$label)

##
visNetwork(nodes_2011, edges_2011, background = "#bee4fb")%>% visOptions(highlightNearest = T)%>%
    visEdges(physics = F, smooth = F, color = "#082e45", shadow = T, width = 2) %>% visIgraphLayout(layout = "layout.davidson.harel") %>% visNodes(labelHighlightBold = T, label = nodes_2011$label, level = 0, font = '14px arial red', color = "#1DA1F2" , shadow =c(T), size = 45, borderWidthSelected = 2) 

###
visNetwork(nodes_2011, edges_2011, background = "#bee4fb")%>% visOptions(highlightNearest = T)%>%
    visEdges(physics = F, smooth = F, color = "#082e45", shadow = T, width = 2, ) %>% visIgraphLayout(layout = "layout.davidson.harel") %>% visNodes(labelHighlightBold = T, label = nodes_2011$label, level = 0, color = "#1DA1F2" , shadow =T, size = 45, borderWidthSelected = 2) #bee4fb


%>%
  visInteraction(hover = T) %>%
  visEvents(hoverEdge = "function(e){
    this.body.data.edges.update({id: e.edge, font: {size : 14}});
  }") %>%
  visEvents(blurEdge = "function(e){
    this.body.data.edges.update({id: e.edge, font: {size : 0}});
  }")

###

visNetwork(nodes_2011, edges_2011, background = "#bee4fb")%>% visOptions(highlightNearest = T)%>%
    visEdges(physics = F, smooth = F, color = "#082e45", shadow = T, width = 2, ) %>% visIgraphLayout(layout = "layout.davidson.harel") %>% 
    visNodes(labelHighlightBold = T, level = 0, color = list(background ="#1DA1F2", highlight = "#91d5ff") , shadow =T, size = 45, borderWidthSelected = 2, font = '40px arial white') %>%
    visInteraction(hover = T) %>% 
    visEvents(blurNode  = "function(e){
            var label_info = this.body.data.nodes.get({
            fields: ['label', 'label_long'],
            filter: function (item) {
            return item.id === e.node
            },
            returnType :'Array'
            });
            this.body.data.nodes.update({id: e.node, label : label_info[0].label_long, label_long : label_info[0].label});
  }")

###

visEvents(selectNode  = "function(e){
            var label_info = this.body.data.nodes.get({
            fields: ['label', 'label_long'],
            filter: function (item) {
            return item.id === e.node
            },
            returnType :'Array'
            });
            this.body.data.nodes.update({id: e.node, label : label_info[0].label_long, label_long : label_info[0].label});
            }") %>% 
  visEvents(blurNode  = "function(e){
            var label_info = this.body.data.nodes.get({
            fields: ['label', 'label_long'],
            filter: function (item) {
            return item.id === e.node
            },
            returnType :'Array'
            });
            this.body.data.nodes.update({id: e.node, label : label_info[0].label_long, label_long : label_info[0].label});
  }"


  
  
  output$hashtag_network_interactive <- renderVisNetwork({
    visNetwork(networkYearInput_interactive()$nodes, networkYearInput_interactive()$edges, background = "#5BBCD6",
               main = list(paste(input$year_interactive, " interactive hashtags network"), "white"), submain = "Unconnected nodes not included")%>% visOptions(highlightNearest = T)%>%
      visEdges(physics = F, smooth = F, color = "#082e45", shadow = T, width = 2 ) %>% visNetwork::visIgraphLayout("layout.davidson.harel") %>% #layout.davidson.harel
      visNodes(labelHighlightBold = T, level = 0, size = input$nodes_size, color = list(background ="#F2AD00", highlight = "#F98400") , shadow =T, borderWidthSelected = 2, 
               font = list(size = input$nodes_size, color =  "white")) %>%
      visInteraction(hover = T, navigationButtons = TRUE) %>%  
      visEvents(blurNode  = "function(e){
          var label_info = this.body.data.nodes.get({
          fields: ['label', 'label_long'],
          filter: function (item) {
          return item.id === e.node
          },
          returnType :'Array'
          });
          this.body.data.nodes.update({id: e.node, label : label_info[0].label_long, label_long : label_info[0].label});
                }")
  
  
```

%>%
    visEvents(selectNode  = "function(e){
            var label_info = this.body.data.nodes.get({
            fields: ['label', 'label_long'],
            filter: function (item) {
            return item.id === e.node
            },
            returnType :'Array'
            });
            this.body.data.nodes.update({id: e.node, label : label_info[0].label_long, label_long : label_info[0].label});
            }") %>% 


```{r}
set.seed(2105)
hashnet_2011 <- ggnetwork(x_2011, layout = "fruchtermanreingold", cell.jitter = 1)
hashnet_2021 <- ggnetwork(x_2021, layout = "fruchtermanreingold", cell.jitter = 0.1)
```


```{r}
#2011
ggplot(hashnet_2011, aes(x, y, xend = xend, yend = yend)) +
     geom_edges(alpha = 0.5) +
     geom_nodes(size = 15, color = "black") +
     geom_nodetext(aes(label = vertex.names), fontface = "bold", color = 'white') +
    theme_blank() +
  theme(panel.background = element_rect(fill = "gray"))


ggraph(hashtags_2011, layout = "kk") +
     geom_edges(alpha = 0.5) +
     geom_nodes(size = 15, color = "black") +
     geom_nodetext(aes(label = vertex.names), fontface = "bold", color = 'white') +
    theme_blank() +
  theme(panel.background = element_rect(fill = "gray"))
```


```{r}
#2021
ggplot(hashnet_2021, aes(x, y, xend = xend, yend = yend)) +
     geom_edges(alpha = 0.5) +
     geom_nodes(size = 15, color = "black") +
     geom_nodetext(aes(label = vertex.names), fontface = "bold", color = "white") +
    theme_blank() + 
  theme(panel.background = element_rect(fill = "gray"))
```


```{r}
x <- as.matrix(hash_2011)
degree_x <- graph.adjacency(x, mode ="undirected", weighted = NULL)


hash$degree = igraph::degree(degree_x)
hash$edge <- E(degree_x)
hash$vertex <- V(degree_x)
# to count edges: gsize(degree_x)
# to count verticies: gorder(degree_x)

# Attach attributes to the vertices of the graph
#V(degree_x)$age = attr$AGE[match(V(degree_x)$name, attr$ID)]
#V(degree_x)$tenure = attr$TENURE[match(V(degree_x)$name, attr$ID)]
#V(degree_x)$level = attr$LEVEL[match(V(degree_x)$name, attr$ID)]
#V(degree_x)$dept = attr$DEPT[match(V(degree_x)$name, attr$ID)]
```

```{r}
simplif

set.seed(2105)

net_df <- ggnetwork(x, 
                    layout = "fruchtermanreingold", 
                    cell.jitter = 0.75)
```

## Visualizations to include
1) Entire 2011 hashtag network
  - options to include/remove unconnected nodes
2) Entire 2021 hashtag network
  - option to include/remove unconnected nodes
3) Option to see 2011-2021 networks
  - option to include/remove unconnected nodes
4) Any grouping algortihm? Girvan-Neuman or Random Walk?
5) Option to simplify(network): simplified networks don't contain loops and multiple edges

# Things to keep in mind
1) color scheme
2) order
3) context / story

                tabPanel("Visualization Explanation",
                         fludrow(box(htmlOutput("network_expl"))))
                         
                         
                           output$network_expl <- renderUI({
    HTML(paste("<h2> Network Analysis </h2> <br>
               ", sep ="<br/>"))
  })
